Evolving Dynamics: The U.S. Labor Force from 1972 to 2022 and the Enduring Challenges for Women in the Workplace

### 2024-03-13 paper/paper.qmd
added results sections for demographic variables

## Respondent Demograpgics

@@

@@

```{r}
#| label: fig-parttime
#| fig-cap: Unemployed Work Status by Gender Over Time
#| echo: false
#| warning: false
#| message: false

@@

ggplot(data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```
## Biases against hiring women

@@

# Results
```{r}
#| label: fig-fehire
#| fig-cap: Attitudes Towards Hiring and Promoting Women
#| echo: false
#| warning: false
#| message: false

@@

## Full-Time Work for Women
@fig-eduTojob reveals a significant upward trend in the number of women engaged in full-time employment over the years. Starting from 191 women in 1972, the numbers generally increased, peaking at 963 women in full-time work in 2006, before experiencing fluctuations and settling at 728 in 2022. This progression underscores the growing integration of women into the workforce and the breaking down of traditional gender roles that confined women to domestic responsibilities. The notable peak in 2006 could be indicative of economic, social, and legislative changes that encouraged or necessitated higher female participation in the workforce.

## Educational Attainment by Gender
Parallel to the rise in full-time employment, @fig-eduTojob also highlights significant strides in women's educational attainment. The data tracks the number of women who have achieved bachelor's and graduate degrees from 1972 through 2022. In 1972, 45 women held bachelor's degrees and 10 held graduate degrees. By 2022, these numbers had surged to 353 for bachelor's degrees and 228 for graduate degrees, reflecting a substantial increase in higher education attainment among women.

>>>>

## Respondent Demographics

@@

@fig-fulltime shows the  number of men and women working full-time from 1972 to 2022, revealing significant trends in labor force participation by gender over five decades. Initially, the workforce was predominantly male, with 570 men working full time compared to 191 women in 1972. Over the years, a steady increase in the number of women working full-time is observed, narrowing the gender gap in full-time employment. By 2022, the number of women working full-time had increased to 728 from 191 in 1972, showing substantial progress towards gender equality in the workplace. The most remarkable year-over-year growth was seen between 1993 and 1994, and again significantly in 2006, reflecting possibly societal, policy, and economic changes that encouraged or necessitated increased female workforce participation.

@@

@fig-part_time highlights the number of men and women engaged in part-time work from 1972 to 2022, showcasing evolving trends in part-time employment, with a noticeable gender dimension. In 1972, part-time work was more common among women (73) than men (41), a trend that persisted and widened over the decades. The gap between male and female part-time workers expanded, with women consistently outnumbering men in part-time positions. Notably, the data illustrates a significant jump in both men and women working part-time in 1994, suggesting a shift in employment patterns or economic conditions that increased part-time work availability or necessity. By 2022, the number of women working part-time reached 219, compared to 126 men, reflecting ongoing gender disparities in part-time employment.

```{r}
#| label: fig-unemployed
#| fig-cap: Unemployed Work Status by Gender Over Time
#| echo: false
#| warning: false
#| message: false

@@

ggplot(data) +
  geom_line(aes(x = year, y = men, color = "Men")) +
  geom_line(aes(x = year, y = women, color = "Women")) +
  scale_color_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(
       x = "Year",
       y = "Number of People",
       color = "Gender") +
  theme_minimal()
```
@fig-unemployed captures the number of unemployed men and women from 1972 to 2022, providing insightful trends into the dynamics of unemployment across genders over five decades. In 1972, the dataset begins with 42 unemployed men compared to 8 unemployed women, indicating a higher male unemployment in absolute numbers. Over the years, both genders experienced fluctuations in unemployment, with notable increases during economic downturns. A significant spike in unemployment for both genders is observed in 2021, with 181 men and 150 women reported as unemployed, possibly reflecting the economic impacts of global events or shifts in the labor market. While the gap between unemployed men and women has varied, the year 2021 shows a notably smaller difference, suggesting a convergence in unemployment rates between genders.

## Biases against hiring women

@@

```{r}
#| label: fig-fehire
#| fig-cap: Attitudes Towards Hiring and Promoting Women
#| echo: false
#| warning: false
#| message: false

@@

## Educational Attainment and Full-Time Work
@fig-eduTojob reveals a significant upward trend in the number of women engaged in full-time employment over the years. Starting from 191 women in 1972, the numbers generally increased, peaking at 963 women in full-time work in 2006, before experiencing fluctuations and settling at 728 in 2022. This progression underscores the growing integration of women into the workforce and the breaking down of traditional gender roles that confined women to domestic responsibilities. The notable peak in 2006 could be indicative of economic, social, and legislative changes that encouraged or necessitated higher female participation in the workforce.

Parallel to the rise in full-time employment, @fig-eduTojob also highlights significant strides in women's educational attainment. The data tracks the number of women who have achieved bachelor's and graduate degrees from 1972 through 2022. In 1972, 45 women held bachelor's degrees and 10 held graduate degrees. By 2022, these numbers had surged to 353 for bachelor's degrees and 228 for graduate degrees, reflecting a substantial increase in higher education attainment among women.
